---
title: "Lab 5: Confidence Interval"
author: "Mamudat Jaji"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(infer)

set.seed(6321)

us_adults <- tibble(
  climate_change_affects = c(rep("Yes", 62000), rep("No", 38000))
)

ggplot(us_adults, aes(x = climate_change_affects)) +
  geom_bar() +
  labs(
    x = "", y = "",
    title = "Do you think climate change is affecting your local community?"
  ) +
  coord_flip() 

us_adults %>%
  count(climate_change_affects) %>%
  mutate(p = n /sum(n))

n <- 60
samp <- us_adults %>%
  sample_n(size = n)

samp %>%
  specify(response = climate_change_affects, success = "Yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
```

### More Practice Exercise 1

Choose a different confidence level than 95%. Would you expect a confidence interval at this level to me wider or narrower than the confidence interval you calculated at the 95% confidence level? Explain your reasoning.

### Answer
I choose a new confidence level of 99% , I would expect a wider confidence interval than that of the 95% confidence level since 99% is more than 95%, it expands and covers more areas than 95%. The higher the confidence level, the wider the confidence interval.

```{r}
samp %>%
  specify(response = climate_change_affects, success = "Yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.99)
```


### More Practice Exercise 2

Using code from the infer package and data from the one sample you have (samp), find a confidence interval for the proportion of US Adults who think climate change is affecting their local community with a confidence level of your choosing (other than 95%) and interpret it.

### Answer
I chose 93% confidence level.
Interpretation: We are 93% confident that 47% to 70% of US adults think climate change is affecting their local community.

```{r}
samp %>%
  specify(response = climate_change_affects, success = "Yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.93)
```


### More Practice Exercise 3

Using the app, calculate 50 confidence intervals at the confidence level you chose in the previous question, and plot all intervals on one plot, and calculate the proportion of intervals that include the true population proportion. How does this percentage compare to the confidence level selected for the intervals?

### Answer
Using the app, I chose 93% confidence level and 50 confidence interval, about 44/50 or 0.88(88%) proportion of intervals were included in the true population. Although, the confidence interval kept changing when I kept running it.

Upon comparison to the confidence level of 95%, the confidence level of 93% is narrower. The confidence interval becomes more narrow as confidence level decrease.

```{r}
# Insert code for Exercise 3 here
```


### More Practice Exercise 4

Lastly, try one more (different) confidence level. First, state how you expect the width of this interval to compare to previous ones you calculated. Then, calculate the bounds of the interval using the infer package and data from samp and interpret it. Finally, use the app to generate many intervals and calculate the proportion of intervals that are capture the true population proportion.

### Answer
I will be using 88% confidence level, I would expect the width to have a narrower confidence interval.

Interpretation: We are 88% confident that 48% to 67% of US adults think climate change is affecting their local community.

Using the app, I chose 88% confidence level and 50 confidence interval, about 45/50 or 0.90(90%) proportion of intervals were included in the true population. Although, the confidence interval kept changing when I kept running it.


```{r}
samp %>%
  specify(response = climate_change_affects, success = "Yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.88)
```


### More Practice Exercise 5

Using the app, experiment with different sample sizes and comment on how the widths of intervals change as sample size changes (increases and decreases).

### Answer

After using different sample sizes, I found out that as the sample size increased, the widths of the confidence interval decreased.

```{r}
# Insert code for Exercise 5 here
```


### More Practice Exercise 6

Finally, given a sample size (say, 60), how does the width of the interval change as you increase the number of bootstrap samples. Hint: Does changing the number of bootstap samples affect the standard error?

### Answer

Using the app, as the number of bootstrap samples increased,the width of the confidence interval decreased. Changing the number of bootstrap samples affects the standard error because it decreases.

```{r}
# Insert code for Exercise 6 here
```
