---
title: "Lab 1: Intro to R"
author: "Mamudat Jaji"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
head(fastfood)

mcdonalds <- fastfood %>%
  filter(restaurant == "Mcdonalds")
dairy_queen <- fastfood %>%
  filter(restaurant == "Dairy Queen")
```

### Exercise 1

In comparison to the dairy queen histogram, the Mcdonalds histogram has a stronger right skew. The dairy queen histogram falls somewhat towards the middle, while the McDonalds cal fat data has a center that is further to the left. The McDonald's data has a wider range, with a maximum of 1200-1400, but the dairy queen data has a maximum of 600-700.

```{r}
# code to show the distribution for calories from fat of mcdonald
hist(mcdonalds$cal_fat)

# code to show the distribution for calories from fat of dairy_queen
hist(dairy_queen$cal_fat)

```


### Exercise 2

Yes, Dairy queen follows a nearly normal distribution.
while Mcdonalds is a rightly skewed distribution.

```{r}
# Insert code for Exercise 2 here
dqmean <- mean(dairy_queen$cal_fat)
dqsd   <- sd(dairy_queen$cal_fat)

ggplot(data = dairy_queen, aes(x = cal_fat)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = dqmean, sd = dqsd), col = "tomato")

```


### Exercise 3

The distribution of this plot appears to be more normal than the original data; it is primarily a diagonal line.

```{r}
# Insert code for Exercise 3 here
ggplot(data = dairy_queen, aes(sample = cal_fat)) + 
  geom_line(stat = "qq")

sim_norm <- rnorm(n = nrow(dairy_queen), mean = dqmean, sd = dqsd)

ggplot(data = dairy_queen, aes(sample = sim_norm)) + 
  geom_line(stat = "qq")

hist(sim_norm)
hist(dairy_queen$cal_fat)
qqnorm(sim_norm)
```


### Exercise 4

The simulated plots appear to have a more evenly-distributed distribution than the data plot.It shows that the calories from fat data for dairy queen is generally a normally distribution.

```{r}
# Insert code for Exercise 4 here

qqnormsim(sample = cal_fat, data = dairy_queen)
```


### Exercise 5

Based on the simulations, the theoretical plots appear to be normally distributed, with a persistent diagonal line on the plots.

```{r}
# Insert code for Exercise 5 here

mdmean <- mean(mcdonalds$cal_fat)
mdsd   <- sd(mcdonalds$cal_fat)

ggplot(data = mcdonalds, aes(x = cal_fat)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = mdmean, sd = mdsd), col = "green")

sim_norm2 <- rnorm(n = nrow(mcdonalds), mean = mdmean, sd = mdsd)

ggplot(data = mcdonalds, aes(sample = sim_norm2)) + 
  geom_line(stat = "qq")

qqnormsim(sample = cal_fat, data = mcdonalds)
```


### Exercise 6

Insert any text here.

```{r}
# Insert code for Exercise 6 here
```


### Exercise 7

Upon comparison, Arby's and Burger King's distribution were closest to normal, however Arby's distribution had the closest to normal for sodium.

```{r}
# Insert code for Exercise 7 here

# creating a code for mean and standard deviation for sodium of Mcdonalds
md_sod_mean <- mean(mcdonalds$sodium)
md_sod_sd   <- sd(mcdonalds$sodium)

# distribution for sodium of Mcdonalds
hist(mcdonalds$sodium)
ggplot(data = mcdonalds, aes(x = sodium)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = md_sod_mean, sd = md_sod_sd), col = "chocolate4")

sim_norm3 <- rnorm(n = nrow(mcdonalds), mean = md_sod_mean, sd = md_sod_sd)

ggplot(data = mcdonalds, aes(sample = sim_norm3)) + 
  geom_line(stat = "qq")

qqnormsim(sample = sodium, data = mcdonalds)

# creating a code for mean and standard deviation for sodium of Dairy Queen
dq_sod_mean <- mean(dairy_queen$sodium)
dq_sod_sd   <- sd(dairy_queen$sodium)

# distribution for sodium of Dairy Queen
hist(dairy_queen$sodium)
ggplot(data = dairy_queen, aes(x = sodium)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = dq_sod_mean, sd = dq_sod_sd), col = "cyan")

sim_norm4 <- rnorm(n = nrow(dairy_queen), mean = dq_sod_mean, sd = dq_sod_sd)

ggplot(data = dairy_queen, aes(sample = sim_norm4)) + 
  geom_line(stat = "qq")

qqnormsim(sample = sodium, data = dairy_queen)

#code to filter Chick Fil A from the dataset
chick_fil_a <- fastfood %>%
  filter(restaurant == "Chick Fil-A")

# creating a code for mean and standard deviation for sodium of Chick Fil A
cfa_sod_mean <- mean(chick_fil_a$sodium)
cfa_sod_sd   <- sd(chick_fil_a$sodium)

# distribution for sodium of Chick Fil-A
hist(chick_fil_a$sodium)
ggplot(data = chick_fil_a, aes(x = sodium)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = cfa_sod_mean, sd = cfa_sod_sd), col = "purple")

sim_norm_cfa_sod <- rnorm(n = nrow(chick_fil_a), mean = cfa_sod_mean, sd = cfa_sod_sd)

ggplot(data = chick_fil_a, aes(sample = sim_norm_cfa_sod)) + 
  geom_line(stat = "qq")

qqnormsim(sample = sodium, data = chick_fil_a)


#code to filter Sonic from the dataset
sonic <- fastfood %>%
  filter(restaurant == "Sonic")

# creating a code for mean and standard deviation for sodium of Sonic
sonic_sod_mean <- mean(sonic$sodium)
sonic_sod_sd   <- sd(sonic$sodium)

# distribution for sodium of Sonic
hist(sonic$sodium)
ggplot(data = sonic, aes(x = sodium)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = sonic_sod_mean, sd = sonic_sod_sd), col = "magenta")

sim_norm_snc_sod <- rnorm(n = nrow(sonic), mean = sonic_sod_mean, sd = sonic_sod_sd)

ggplot(data = sonic, aes(sample = sim_norm_snc_sod)) + 
  geom_line(stat = "qq")

qqnormsim(sample = sodium, data = sonic)

#code to filter Arbys from the dataset
arb <- fastfood %>%
  filter(restaurant == "Arbys")

# creating a code for mean and standard deviation for sodium of Arbys
arb_sod_mean <- mean(arb$sodium)
arb_sod_sd   <- sd(arb$sodium)

# distribution for sodium of Arbys
hist(arb$sodium)
ggplot(data = arb, aes(x = sodium)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = arb_sod_mean, sd = arb_sod_sd), col = "darkcyan")

sim_norm_arb_sod <- rnorm(n = nrow(arb), mean = arb_sod_mean, sd = arb_sod_sd)

ggplot(data = arb, aes(sample = sim_norm_arb_sod)) + 
  geom_line(stat = "qq")

qqnormsim(sample = sodium, data = arb)

#code to filter Burger King from the dataset
bgk <- fastfood %>%
  filter(restaurant == "Burger King")

# creating a code for mean and standard deviation for sodium of Burger King
bgk_sod_mean <- mean(bgk$sodium)
bgk_sod_sd   <- sd(bgk$sodium)

# distribution for sodium of Burger King
hist(bgk$sodium)
ggplot(data = bgk, aes(x = sodium)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = bgk_sod_mean, sd = bgk_sod_sd), col = "darkblue")

sim_norm_bgk_sod <- rnorm(n = nrow(bgk), mean = bgk_sod_mean, sd = bgk_sod_sd)

ggplot(data = bgk, aes(sample = sim_norm_bgk_sod)) + 
  geom_line(stat = "qq")

qqnormsim(sample = sodium, data = bgk)

#code to filter Subway from the dataset
swy <- fastfood %>%
  filter(restaurant == "Subway")

# creating a code for mean and standard deviation for sodium of Subway
swy_sod_mean <- mean(swy$sodium)
swy_sod_sd   <- sd(swy$sodium)

# distribution for sodium of Subway
hist(swy$sodium)
ggplot(data = swy, aes(x = sodium)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = swy_sod_mean, sd = swy_sod_sd), col = "coral")

sim_norm_swy_sod <- rnorm(n = nrow(swy), mean = swy_sod_mean, sd = swy_sod_sd)

ggplot(data = swy, aes(sample = sim_norm_swy_sod)) + 
  geom_line(stat = "qq")

qqnormsim(sample = sodium, data = swy)

#code to filter Taco Bell from the dataset
tcb <- fastfood %>%
  filter(restaurant == "Taco Bell")

# creating a code for mean and standard deviation for sodium of Taco Bell
tcb_sod_mean <- mean(tcb$sodium)
tcb_sod_sd   <- sd(tcb$sodium)

# distribution for sodium of Taco Bell
hist(tcb$sodium)
ggplot(data = tcb, aes(x = sodium)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = tcb_sod_mean, sd = tcb_sod_sd), col = "rosybrown")

sim_norm_tcb_sod <- rnorm(n = nrow(tcb), mean = tcb_sod_mean, sd = tcb_sod_sd)

ggplot(data = tcb, aes(sample = sim_norm_tcb_sod)) + 
  geom_line(stat = "qq")

qqnormsim(sample = sodium, data = tcb)
```

### Exercise 8

Step wise patterns occurs in distributions that are integers and discrete, these might be the case since these variables are integers and discrete.

```{r}

```

### Exercise 9

According to the plots, the distribution of total carbs for Arby's is nearly symmetrical.

```{r}

arb_carb_mean <- mean(arb$total_carb)
arb_carb_sd   <- sd(arb$total_carb)

ggplot(data = arb, aes(x = total_carb)) +
        geom_blank() +
        geom_histogram(aes(y = ..density..), bins = 18) +
        stat_function(fun = dnorm, args = c(mean = arb_carb_mean, sd = arb_carb_sd), col = "violetred")

sim_norm_arb_carb <- rnorm(n = nrow(arb), mean = arb_carb_mean, sd = arb_carb_sd)

ggplot(data = arb, aes(sample = sim_norm_arb_carb)) + 
  geom_line(stat = "qq")

hist(arb$total_carb)

```

