---
title: "Lab 3: Probability"
author: "Olorunleke"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)

# Setting a seed
set.seed(9500) 

# to view and load the data
glimpse(kobe_basket)

# to add the data set
kobe_basket <- kobe_basket
```

### Exercise 1

What does a streak length of 1 mean, i.e. how many hits and misses are in a streak of 1? What about a streak length of 0?

The streak length of 1 is 24 while the streak length of 0 is 39.

```{r}

# to subset and create a data set for the streak length
kobe_streak <- calc_streak(kobe_basket$shot)

# to create a bar chart of the streak length
ggplot(data = kobe_streak, aes(x = length)) +
  geom_bar()
```


### Exercise 2

Describe the distribution of Kobe’s streak lengths from the 2009 NBA finals. What was his typical streak length? How long was his longest streak of baskets? Make sure to include the accompanying plot in your answer.

His typical streak length according to the bar chart is 0. His longest basket streak is 4.



```{r}
# Insert code for Exercise 2 here

# to create a bar chart of the streak length
ggplot(data = kobe_streak, aes(x = length)) +
  geom_bar()
```


### Exercise 3

Insert any text here.

```{r}
# Insert code for Exercise 3 here

coin_outcomes <- c("heads", "tails")
sample(coin_outcomes, size = 1, replace = TRUE)

coin_outcomes <- c("heads", "tails")
sample(coin_outcomes, size = 100, replace = TRUE)

sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)

sim_fair_coin
table(sim_fair_coin)

sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))

sim_unfair_coin
table(sim_unfair_coin)
```


### Exercise 4

Insert any text here.

```{r}
# Insert code for Exercise 4 here

# simulating an independent shooter
shot_outcomes <- c("H", "M")
 sample(shot_outcomes, size = 1, replace = TRUE)

 # simulating a new independent shooter with a probability of 45% for 133 shots
sim_basket <- sample(shot_outcomes, size = 133, replace = TRUE, prob = c(0.45, 0.55))

# to tabulate sim_basket
sim_basket
table(sim_basket)
```


### Exercise 5

Insert any text here.

```{r}
# Insert code for Exercise 5 here

# subsetting a new data set to show the computation for streak length of sim basket
sim_streak <- calc_streak(sim_basket)

```


### Exercise 6

Describe the distribution of streak lengths. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? How long is the player’s longest streak of baskets in 133 shots? Make sure to include a plot in your answer.

The typical streak length is 0 and the longest streak length is 5.

```{r}
# Insert code for Exercise 6 here

# plotting a bar graph to show the distribution for sim_streak
ggplot(data = sim_streak, aes(x = length)) +
  geom_bar()
```


### Exercise 7

If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.

After running the simulation again, it came up as 63 70, while the previous one was 52 81, so it is totally different. The hit became greater while the miss became less.

```{r find-max-total}
# Insert code for Exercise 7 here

 # re-running simulation a new independent shooter with a probability of 45% for 133 shots
sim_basket <- sample(shot_outcomes, size = 133, replace = TRUE, prob = c(0.45, 0.55))

# to tabulate sim_basket
sim_basket
table(sim_basket)
```

### Exercise 8

Both distribution are quite similar, the most streak was 0, the second most was 1. Although, the number of bins for Kobe Bryant's streak is 5 while for the independent shooter is 6.

The hot hand model does not fit Kobe Bryant’s shooting patterns because both Kobe and the independent shooter does not have high of success and their longest number streaks had shorter lenght

```{r}

# to create a bar chart of the streak length for Kobe
ggplot(data = kobe_streak, aes(x = length)) +
  geom_bar()

# to create a bar chart of the streak length for the independent shooter
ggplot(data = sim_streak, aes(x = length)) +
  geom_bar()
```

