---
title: "Lab 1: Intro to R"
author: "Olorunleke"
date: "`r Sys.Date()`"
output: pdf_document
editor_options:
  chunk_output_type: inline
---
arbuthnot
install.packages("tidyverse")
tinytex::install_tinytex()
--- 
```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

Insert any text here.

```{r trend-girls}
# Insert code for Exercise 2 here
```


### Exercise 3

Insert any text here.

```{r plot-prop-boys-arbuthnot}
# Insert code for Exercise 3 here
```


### Exercise 4

The year 1940-2002. 63 observations of 3 variables. Year, boys and girls.

```{r dim-present}
# code to shows the data set
present
present <- present

# these code shows the years, dimension and variable names in the data set
present$year

dim(present)

names(present)
```


### Exercise 5

Upon comparison,only the variable names are similar, the dimension and year are on a different scale.

```{r count-compare}
# these code shows the year range in present and arbuthnot data set
range(present$year)
range(arbuthnot$year)

# these code shows the dimension in present and arbuthnot data set
dim(present)
dim(arbuthnot)

# these code shows the variable names in present and arbuthnot data set
names(present)
names(arbuthnot)
```


### Exercise 6

Arburthnot's observation holds because the proportion of boys born overtime is greater than 0.5, while for girls it is less than 0.5.

```{r plot-prop-boys-present}
# This code adds the variable 'total' to present data set
  present <- present %>%
  mutate(total = boys + girls)
  
# This code shows the proportion of boys overtime and adds the variable to the data set
  present <- present %>%
  mutate(boy_ratio = boys / total)
  
# This code shows the plot the proportion of boys
ggplot(data = present, aes(x= year, y = boy_ratio)) + geom_line()

# This code shows the proportion of girls overtime and adds the variable to the data set
present <- present %>%
  mutate(girl_ratio = girls / total)
  
# This code shows the plot for the proportion of girls
  ggplot(data = present, aes(x= year, y = girl_ratio)) + geom_line()
  
   present <- present %>%
  mutate(more_boys = boys > girls)
```


### Exercise 7

The year 1961 has the highest total number of births with 4268326.

```{r find-max-total}
# This code adds the variable total to the data set
 present <- present %>%
  mutate(total = boys + girls)

# This code arranges the data set in descending order
  present %>%
  arrange(desc(total))
  
```

